---
title: 'Geog4/6300: Lab 0'
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
```

*Your name: {Your name here}*

This "lab" assignment provides an opportunity to learn the basic mechanics of Github classroom, loading data, and doing some basic operations using the tools in the tidyverse. 

We'll be combining two datasets for this assignment: the New York Times dataset on COVID case counts by county and American Community Survey data on county population. You can load them both using the read_csv function and either the URL or path. For example, you can load the ACS data using this function:

```{r}
county_pop<-read_csv("data/pop_acs18.csv")
```

In this dataset, you'll see a county FIPS code (GEOID), county name, the variable code for total population, a population estimate, and a margin of error (which is blank in this case). 

## Loading the data

### Question 1:
Load these data using the read_csv command. You can copy the code above for the county population data. For the New York Times data, replace the file path name with this URL:

* https://github.com/nytimes/covid-19-data/raw/master/us-counties.csv

In addition, filter the data so you only have data for May 13, 2022, the last date of observations in the dataset.

```{r}
##Your code goes here.
```

## Connecting the data

Next you'll need to join these two datasets using the inner_join function. To do so you'll need to have two fields with the same name in each dataset. The fips codes are in both datasets, but they have different names (GEOID and fips) and formats (numeric and character). The following code creates a new variable called fips in the ACS data that's a character using the mutate function.

```{r}
county_pop<-county_pop %>%
  mutate(fips=as.character(GEOID))
```

Call the function above to rename your data. Now you're ready to join the data.

### Question 2
Using inner_join, join the New York Times dataset to the county population one. 

```{r}

```

### Question 3
Find some documentation on the inner_join function online or using help in R. Describe how it works, and explain how the results would have been different if you used full_join instead.

{Write your response here.}

## Calculating a rate
Now you can calculate a case rate by county, also using mutate. These are usually reported as cases per 100,000 population You'll need to set up some math that follows the following pattern:

* newvar = cases / population * 100000

### Question 4
Use mutate to calculate a case rate per 100,000 people.

```{r}
#Your code goes here
```

### Question 5
Which Georgia county has the *highest* case rate based on total cases according to your analysis? You can open the dataset and click on the column name for the rate to arrange from high to low. Give the county name, the rate, and the date when this rate was calculated.

{Your answer goes here.}


### Challenge question
What if you wanted to study only those cases happening in the last six months of the dataset, from November 13, 2021 until May 13, 2022? Use the tools you have been learning for data wrangling in class to identify the number of cases per 100,000 residents during that time period in Georgia. Show your code and call the head of the table when done.

```{r}
#Your code goes here
```

### Question 6
Having completed this lab, how do you feel you did on this assignment? Did you hit a home run or just scrape by? What was difficult and/or easy? What did you learn?

{Your response goes here.}

That's it! When you're done with this lab, use the Knit command to render it as a Github markdown document. Then push it up to Github using the procedure outlined in this week's videos.